---
title: "Senior Thesis"
author: "Tung Nguyen"
date: "2/9/2020"
output:
  md_document: default
  pdf_document: default
  word_document: default
  html_document: default
---
```{r include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo=FALSE, comment=FALSE, include=FALSE)
```

```{r Loading packages, warning=FALSE}
# library(foreign) # read in sav file
# install.packages("magrittr") # package installations are only needed the first time you use it
# install.packages("dplyr")    # alternative installation of the %>%
library(magrittr) # needs to be run every time you start R and want to use %>%
#install.packages('testit') # assert function
library(testit)
library(data.table)
#install.packages('stringr')
library(glue)

```


# Generate Data for Simulation Experiment


```{r Generate Data for Simulation Experiment}
# knitr::opts_chunk$set(echo=FALSE, warning=TRUE)

# Simulate data to test the power of the statistical test
## ----
# new_data = matrix(0, nrow=500, ncol=6)

# Generate marginal probabilities for W_i
pi_R = c(0.4,0.6)
fun = function(p)((p/(0.4-p))/((0.5-p)/(0.1+p)) - 2)
pw.11 = uniroot(fun, lower=0.0, upper = 1.0)$root
pw.10 = 0.4 - pw.11
pw.01 = 0.5 - pw.11
pw.00 = 0.1 + pw.11
pw.11 + pw.10 + pw.01 + pw.00

# Generate marginal probabilities for Y_i
pi_C = c(0.2, 0.3)
py.11 = 0.085
py.10 = 0.2 - py.11
py.01 = pi_C[2] - py.11
py.00 = 0.5 + py.11
py.11 + py.00 + py.10 + py.01
# shuffle W - Alternative therapies probabilities
orderw = cbind(c(11,10,01,00),c(pw.11,pw.10,pw.01,pw.00))
orderw1 = orderw[sample(nrow(orderw)),]
orderw2 = cbind(orderw1, cumsum(orderw1[,2]))

# shuffle Y - Recommendation sources probabilities
ordery = cbind(c(11,10,01,00),c(py.11,py.10,py.01,py.00))
ordery1 = ordery[sample(nrow(ordery)),]
ordery2 = cbind(ordery1, cumsum(ordery1[,2])) 


#-------------------------------------
# Simulate data using computed marginal probabilities

generate_y = function(proby, data, Uy, row, pos1){
  #' @description Generate simulated data for the Y variable using specified probabilities
  #' @param proby matrix. A 1x3 matrix containing cumulative probability and its corresponding cluster
  #' @param data matrix. A 1x6 empty matrix to store data simulated for the Y variables
  #' @param Uy decimal. A decimal number randomly generated from a uniform distribution.
  #' @param pos1 integer. Integer numbers specifying columns to fill in data
  #' @param row integer. Interger number specifying the row to fill in data
  #' @usage generate_y(proby, data, Uy)
  #' @return a 500x6 matrix that contains simulated data from Y
  if (Uy >= proby[1,3] & Uy < proby[2,3]){
    data[row,pos1] = 0
    data[row, pos1+1] = 0
  } else if (Uy >= proby[2,3] & Uy < proby[3,3]){
    data[row, pos1] = 1
    data[row, pos1+1] = 0
  } else if (Uy >= proby[3,3] & Uy < proby[4,3]){
    data[row, pos1] = 1
    data[row, pos1+1] = 1 
  } else {
    data[row, pos1] = 0
    data[row, pos1+1] = 1
  }
  return(data)
}
generate_cluster = function(probw,proby, Uw, Uy1, Uy2, i, pos1, pos2, row, data=NULL){
#' @description Generate simulated data for W using spcified probabilities
#' @param probw,proby matrix. A 1x3 matrix containing cumulative probability and its corresponding cluster
#' @param data matrix. A 500x6 empty matrix to store simulated data
#' @param Uw,Uy1,Uy2 float. A float number generated from a uniform distribution
#' @param pos1,pos2 integer. Numbers specifying the columns to fill in the data with
#' @param i integer. A number specifying the row to fill in the data with
#' @usage generated_cluster(prob, data, U)
#' @return a 500x6 matrix that contains simulated data
if(is.null(data)){
data = matrix(NA, nrow=500, ncol=6)
colnames(data) = c('W1','W2','Y11','Y12','Y21','Y22')
}
  if (Uw >= probw[1,3] & Uw < probw[2,3]){
    data[i,1] = 0
    data[i,2] = 0
    data[i,3:6] = NA
  }
  else if (Uw >= probw[2,3] & Uw < probw[3,3]){
    data[i, 1] = 1
    data[i, 2] = 0
    data = generate_y(proby, data, Uy1,row=i, pos1)
  }
  else if (Uw >= probw[3,3] & Uw < probw[4,3]){
    data[i, 1] = 1
    data[i, 2] = 1
    data = generate_y(proby, data, Uy1, row=i, pos1)
    data = generate_y(proby, data, Uy2, row=i, pos2)
  }
  else {
   data[i,1] = 0
   data[i,2] = 1
   data = generate_y(proby, data, Uy2, row=i, pos2)
  }
  return (data)
}

# Simulating Data for Experiment
simulate_data = function(filepath, n=500){
count = 1
Uw = runif(1, 0, 1)
Uy1 = runif(1, 0, 1)
Uy2 = runif(1, 0, 1)
simulated_data = generate_cluster(probw = orderw2, proby = ordery2, Uw = Uw, Uy1=Uy1, Uy2=Uy2, i= count, pos1=3, pos2=5)
while (count < n){
  Uw = runif(1, 0, 1)
  Uy1 = runif(1, 0, 1)
  Uy2 = runif(1, 0, 1)
  simulated_data = generate_cluster(probw = orderw2, proby = ordery2, Uw = Uw, Uy1=Uy1, Uy2=Uy2, i= count+1, pos1=3, pos2=5, data=simulated_data)
  count=count+1
  }
  # Change the order of columns to reflect the true order of the data due to the way Y's were generated
  simulated_data[,c(1,2,3,4,5,6)] = simulated_data[,c(1,2,3,5,4,6)]
  # Save simulated data to filepath 
  write.csv(simulated_data, file = filepath, row.names = FALSE)
  return (simulated_data)
}

# Execute create simulated data and save to filepath
no = 1
filepath = glue('./{no}.csv')
print(filepath)
simulated_data = simulate_data(filepath = filepath)
```

```{r functions to calculate test statistics}
get_ij = function(data, cond){
  #' @description Calculate the count of ij
  ij_count = matrix(0, nrow=2, ncol=2)
  ij_count[1,1] = dim(subset(data,(data[,1]==1 & data[,3]==cond)))[1]
  ij_count[1,2] = dim(subset(data,(data[,1]==1 & data[,5]==cond)))[1]
  ij_count[2,1] = dim(subset(data,(data[,2]==1 & data[,4]==cond)))[1]
  ij_count[2,2] = dim(subset(data,(data[,2]==1 & data[,6]==cond)))[1]
return (ij_count)
}
get_i = function(data){
  #' @description Calculate the count of i
  i_count = matrix(0, nrow = 1,ncol = 2)
  i_count[,1] = dim(data[data[,1]==1,])[1]
  i_count[,2] = dim(data[data[,2]==1,])[1]
return (i_count)
}

get_j = function(data, cols, cond, j, j_count = NULL){
  #' @description Calculate the count of j
test_data = subset(data, (data[,cols[1]]==cond | data[,cols[2]]==cond))
if (is.null(j_count)){
j_count = matrix(0, nrow = 1, ncol = 2)
}
for (indi in 1:nrow(test_data)){
  if (is.na(sum(data[indi,cols]))){
   j_count[1,j] = j_count[1,j] + 1
  }
  else {
    #[Solved When cond = 0 , the sum = 0 when it should be 1 or 2
  j_count[1,j] = j_count[1,j] + sum(data[indi,cols]==cond) / 2
  }
}
return (j_count)
}
# Calculate chi-squared statistics
get_stats = function(data, sequence=FALSE){
  n = dim(data)[1]
  ij_count1 = get_ij(data=data, cond= 1)
  notij_count1 = get_ij(data=data, cond= 0)
  i_count1 = get_i(data=data)
  yj_count1 = get_j(data=data, cols=c(5,6), cond=1, j=2)
  yj_count1 = get_j(data=data, cols=c(3,4), cond=1, j=1, j_count=yj_count1)
  notj_count1 = get_j(data=data, cols=c(5,6), cond=0, j=2)
  notj_count1 = get_j(data=data, cols=c(3,4), cond=0, j=1, j_count=notj_count1)
  pi1_i. = i_count1 / n
  # the grand total proportion of yeses and nos for the ith, jth and ijth values
  yp1_.j = yj_count1 / n
  np1_.j = notj_count1 / n
  ypi1_ij = ij_count1 / n
  npi1_ij = notij_count1 / n
  yp_grandstat1 = sum((ypi1_ij - t(pi1_i.) %*% yp1_.j)^2 / (t(pi1_i.) %*% yp1_.j))
  np_grandstat1 = sum((npi1_ij - t(pi1_i.) %*% np1_.j)^2 / (t(pi1_i.) %*% np1_.j))
  grandstat1 = (yp_grandstat1 + np_grandstat1) * n
  if (sequence==TRUE){
    return (list(grandstat = grandstat1, yp_grandstat = yp_grandstat1, np_grandstat = np_grandstat1))
  }
  else {
    return (grandstat1)
}
}

# [TODO]: Sample from only non-nan values for W, Y
cal_p_boot = function(data, X, B.max=1999, FUN=get_stats){
  stat = FUN(data=data)
  X.sq.S.star = numeric(length(B.max))
  counter = 0
  b = 0
  while(b <= B.max){
    b = b+1
    n = dim(data)[1]
    end = dim(data)[2]
    W = sample(x=1:n,size=n,replace=TRUE)
    c3 = sample(data[!is.na(data[,3]),3], size=n, replace = TRUE)
    c4 = sample(data[!is.na(data[,4]),4], size=n, replace = TRUE)
    c5 = sample(data[!is.na(data[,5]),5], size=n, replace = TRUE)
    c6 = sample(data[!is.na(data[,6]),6], size=n, replace = TRUE)
    data.star = cbind(data[W,1:X],c3,c4,c5,c6)
    data.star[data.star[,'W1']==0, c(3,5)] = NA
    data.star[data.star[,'W2']==0, c(4,6)] = NA
    stat.star = FUN(data.star)[1]
    
    counter = counter + 1
    X.sq.S.star[counter] = stat.star
  }
  p.value.boot = mean(X.sq.S.star >= stat)
  return(list(p_value=p.value.boot, X.sq=X.sq.S.star))
}
test = cal_p_boot(simulated_data, X=2)

```

```{r Sample from only non-nan values for W, Y}
data = copy(simulated_data)
B.max=1999
stat = get_stats(data=data)
X.sq.S.star = numeric(length(B.max))
counter = 0
b = 0
n = 500
X = 2
while(b <= B.max){
  b = b+1
  n = dim(data)[1]
  end = dim(data)[2]
  W = sample(x=1:n,size=n,replace=TRUE)
  c3 = sample(data[!is.na(data[,3]),3], size=n, replace = TRUE)
  c4 = sample(data[!is.na(data[,4]),4], size=n, replace = TRUE)
  c5 = sample(data[!is.na(data[,5]),5], size=n, replace = TRUE)
  c6 = sample(data[!is.na(data[,6]),6], size=n, replace = TRUE)
  data.star = cbind(data[W,1:X],c3,c4,c5,c6)
  data.star[data.star[,'W1']==0, c(3,5)] = NA
  data.star[data.star[,'W2']==0, c(4,6)] = NA
  stat.star = get_stats(data.star)[1]

  counter = counter + 1
  X.sq.S.star[counter] = stat.star
  }
  p.value.boot = mean(X.sq.S.star >= stat)
  return(list(p_value=p.value.boot, X.sq=X.sq.S.star))

```


```{r Read in data}
simulated_data = read.csv('1.csv')
data = copy(simulated_data)
n = dim(data)[1]
ij_count1 = get_ij(data=data, cond= 1)
notij_count1 = get_ij(data=data, cond= 0)
i_count1 = get_i(data=data)
yj_count1 = get_j(data=data, cols=c(5,6), cond=1, j=2)
yj_count1 = get_j(data=data, cols=c(3,4), cond=1, j=1, j_count=yj_count1)
notj_count1 = get_j(data=data, cols=c(5,6), cond=0, j=2)
notj_count1 = get_j(data=data, cols=c(3,4), cond=0, j=1, j_count=notj_count1)
pi1_i. = i_count1 / n
# the grand total proportion of yeses and nos for the ith, jth and ijth values
yp1_.j = yj_count1 / n
np1_.j = notj_count1 / n
ypi1_ij = ij_count1 / n
npi1_ij = notij_count1 / n
yp_grandstat1 = sum((ypi1_ij - t(pi1_i.) %*% yp1_.j)^2 / (t(pi1_i.) %*% yp1_.j))
np_grandstat1 = sum((npi1_ij - t(pi1_i.) %*% np1_.j)^2 / (t(pi1_i.) %*% np1_.j))
grandstat1 = (yp_grandstat1 + np_grandstat1) * n
```

```{r Get i,j,ij for data.star}
data = copy(simulated_data)
n = dim(data)[1]
end = dim(data)[2]
W = sample(x=1:n,size=n,replace=TRUE)
Y = sample(x=1:n,size=n,replace = TRUE)
X=2
data.star = cbind(data[W,1:X],data[Y,(X+1):end])
data.star[data.star[,'W1']==0, c(3,5)] = NA
data.star[data.star[,'W2']==0, c(4,6)] = NA
ij_count = get_ij(data=data.star, cond= 1)
notij_count = get_ij(data=data.star, cond= 0)
i_count = get_i(data=data.star)
yj_count = get_j(data=data.star, cols=c(5,6), cond=1, j=2)
yj_count = get_j(data=data.star, cols=c(3,4), cond=1, j=1, j_count=yj_count)
notj_count = get_j(data=data.star, cols=c(5,6), cond=0, j=2)
notj_count = get_j(data=data.star, cols=c(3,4), cond=0, j=1, j_count=notj_count)
pi_i. = i_count / n
# the grand total proportion of yeses and nos for the ith, jth and ijth values
yp_.j = yj_count / n
np_.j = notj_count / n
ypi_ij = ij_count / n
npi_ij = notij_count / n
yp_grandstat = sum((ypi_ij - t(pi_i.) %*% yp_.j)^2 / (t(pi_i.) %*% yp_.j))
np_grandstat = sum((npi_ij - t(pi_i.) %*% np_.j)^2 / (t(pi_i.) %*% np_.j))
grandstat = (yp_grandstat + np_grandstat) * n
```


```{r Get grand stats}
get_stats(data=simulated_data)
```

# Generating bootstrapping data
```{r Generating bootstrapping data, echo=FALSE}
# [TO DO] This is where I can implement parallel programming
# cal_p_boot = function(data){
# X = 2
# n = dim(data)[1]
# end = dim(data)[2]
# W = sample(x=1:n,size=n,replace=TRUE)
# Y = sample(x=1:n,size=n,replace = TRUE)
# data.star = cbind(data[W,1:X],data[Y,(X+1):end])
# # Handling 0 values from W, turning corresponding Y values into NA
# data.star[data.star[,'W1']==0, c(3,5)] = NA
# data.star[data.star[,'W2']==0, c(4,6)] = NA
# get_stats(data.star)
# }
